Mining the human microbiome for clinical insight

### 2019-01-03 chap_24hr.tex
Update with latest versions

\noindent
Mariana G Matus, Claire Duvallet, Melissa Soule, Noriko Endo, Newsha Ghaeli,  Ilana Brito, Carlo Ratti, Elizabeth B Kujawinski, Eric J Alm

>>>>

\noindent
Mariana Matus, Claire Duvallet, Melissa Kido Soule, Krista Longnecker, Noriko Endo, Newsha Ghaeli, Ilana Brito, Carlo Ratti, Elizabeth B Kujawinski, Eric J Alm

### 2019-01-03 donor-selection/ms.tex
Update with latest versions

\begin{figure}
    \begin{center}
    \includegraphics[width=\textwidth]{fig4_power_simulation.png}
    \caption{Power simulation results, showing how many of the 10 most ``truly'' differentially abundant genera would be recovered as significant under different FMT study designs. Each panel represents a different FMT response rate (i.e. percent of patients in the responder vs. non-responder group). The effect size (i.e. number of genera which are differentially abundant in FMT responders vs. non-responders) was simulated by using three different case-control microbiome datasets. A large effect size is modeled by the effect of diarrhea on the microbiome, medium by colorectal cancer, and small by obesity. The top 10 "true" differentially abundant genera were identified by calculating their signal-to-noise ratios in the full original dataset (i.e. mean difference divided by the standard deviation).}\label{fig:power-sim}
    \end{center}
\end{figure}

@@

As clinical trial design methodologies for FMT trials become more developed, many additional questions will need to be addressed.
Many of these key questions relate to the nuances involved in choosing healthy donors: what defines a ``healthy'' donor, and what should define one?
These questions are critical because regardless of the underlying model, in all cases a healthy donor must be identified.
However, as our understanding of the microbiome in societies around the world continues to increase, consensus on the exact structure of a ``healthy'' microbiome decreases.
Should donors be selected to reflect the patient populations, or simply be devoid of pathogens and source from as ``healthy'' of donors as possible?
We know that Europeans and North Americans tend to have less \textit{Prevotella} than healthy Africans from across the continent in both urban and rural communities \cite{Yatsunenko2012,Ou2013,DeFilippo2010}.
Thus, should a clinician carrying out an FMT trial in an African setting use local African cohort as their donor population in order to better match the expected ``healthy'' state of their participants?
In some cases, using a local population may be incompatible with established donor screening criteria because of geographical differences in baseline microbial community composition, prior likelihood of pathogen colonization, or differences in diet and lifestyle habits leading to different dominating strains.
Should clinicians consider relaxing donor screening and exclusion criteria in cases where donors are sourced from countries where commensal colonization by potential pathogens is common population-wide?
More research to understand the functional differences and clinical implications of different ``healthy'' communities across global populations and specifically in the context of FMT trials will be required before these and many related questions can be answered \cite{Bello2018,Rabesandratana2018}.

On the patient side, comorbidities, lifestyle, and dynamic disease manifestations may present challenges and opportunities to improve donor selection processes.
How should comorbidities be incorporated into donor selection?
Patients with multiple disease processes may be dominated by one microbiome-mediated disease model or may exhibit a combination of models, perhaps affecting which donors would be optimal for their specific case.
For example, a person with one community-level function process and one acute dysbiosis process may respond well to total community replacement alone, or may require a combination of total community replacement along with enrichment for community function.
Additionally, will diseases that change manifestations over time benefit from employing different donor selection strategies over the course of a longitudinal FMT trial?
Although there have been no serious adverse events related to FMT material in either clinical practice for rCDI or in clinical trials across adults or pediatrics, could some donors further reduce the probability of adverse events in at-risk patients?
Finally, how should other sources of heterogeneity like lifestyle, diet, and medication usage be incorporated into rational donor selection?
In cases where FMT is combined with other microbiome-targeted interventions like prebiotics or dietary changes, could some donors have synergistic effects with these paired interventions and lead to greater clinical success?

To ensure that FMT reaches its full potential to improve and save patient lives, clinicians should think critically about how their FMT trials can be designed for maximal impact.
By applying new approaches like rational donor selection and adaptive trial designs, the number of trials which fail even though they could have succeeded will drastically decrease.
Furthermore, by developing targeted hypotheses, post-trial analysis plans, and associated sample collection schema alongside the core FMT trial design itself, the number of basic scientific discoveries that are made from each trial will significantly increase.
As FMT expands far beyond rCDI and microbiome-based therapeutics are developed to target a range of disease, novel methods and approaches tailored to the unique challenges and opportunities presented by FMT will be critical to ensuring the advancement of translational microbiome science and beneficial impact on patient lives.

>>>>

\begin{figure}
    \begin{center}
    \includegraphics[width=\textwidth]{fig4_power_simulation.png}
    \caption{Power simulation results, showing how many of the 10 most ``truly'' differentially abundant genera would be recovered as significant under different FMT study designs. Each panel represents a different FMT response rate (i.e. percent of patients in the responder vs. non-responder group). The effect size (i.e. number of genera which are differentially abundant in FMT responders vs. non-responders) was simulated by using three different case-control microbiome datasets. A large effect size is modeled by the effect of diarrhea on the microbiome, medium by colorectal cancer, and small by obesity. The top 10 ``true'' differentially abundant genera were identified by calculating their signal-to-noise ratios in the full original dataset (i.e. mean difference divided by the standard deviation).}\label{fig:power-sim}
    \end{center}
\end{figure}

@@

As FMT-specific clinical trial design methodologies become more developed, many additional questions will need to be addressed.
Some of these key questions relate to choosing healthy donors: what defines a "healthy" donor, and when and how should that definition change?
Some screening criteria, like excluding certain known pathogens, will likely apply to all donors.
Beyond these, however, there is little consensus on what defines a "healthy" microbiome.
Consider a clinician carrying out an FMT trial in an African setting: given that healthy Africans from across the continent are known to have more \textit{Prevotella} than North Americans, it might be advisable to source donors locally to better match the expected healthy state of the patients \cite{Ou2013,Yatsunenko2012,DeFilippo2010}.
But what if the local population has higher asymptomatic colonization rates of undesirable bacteria like opportunistic pathogens, resulting in usual screening criteria excluding many or all potential donors: should the criteria be adapted to allow for local donors, or should donors be sourced from a foreign stool bank whose donors may not match the local population?
To answer this question, more research will need to be carried out to understand differences between healthy microbiomes globally and their clinical implications \cite{Bello2018,Rabesandratana2018}.

On the patient side, comorbidities, lifestyle, and dynamic disease manifestations present additional challenges and opportunities to improve donor selection and FMT clinical trial designs.
How should comorbidities be incorporated into donor selection?
Patients with multiple disease processes may be dominated by one disease model or may exhibit a combination of models, perhaps affecting which donors would be optimal for their specific case.
For example, a person whose condition involves both acute dysbiosis and community-level dysfunction might respond well to any healthy donor, or may require a more complex combination of total community replacement along with enrichment for community function.
Relatedly, diseases that change manifestations over time may benefit from employing different donor selection strategies over the course of a longitudinal FMT trial.
Additionally, although there have been no serious adverse events related to FMT material in either clinical practice for rCDI or in clinical trials across adults or pediatrics, could some donors further reduce the probability of adverse events in at-risk patients?
Finally, how should other sources of heterogeneity like lifestyle, diet, and medication usage be incorporated into rational donor selection?
In cases where FMT is combined with other microbiome-targeted interventions like prebiotics or dietary changes, could some donors have synergistic effects with these paired interventions and lead to greater clinical success?

To ensure that FMT reaches its full potential to improve and save patient lives, clinicians should think critically about how their FMT trials can be designed for maximal impact.
By applying new approaches like rational donor selection and adaptive trial designs, the number of trials which fail even though they could have succeeded will decrease.
Furthermore, by developing targeted hypotheses, post-trial analysis plans, and associated sample collection schema alongside the core FMT trial design itself, the number of basic scientific discoveries that are made from each trial will significantly increase.
As FMT expands beyond rCDI and microbiome-based therapeutics are developed to target a range of diseases, novel methods and approaches tailored to the unique challenges and opportunities presented by FMT will be critical to ensuring the advancement of translational microbiome science and beneficial impact on patient lives.
